---
title: "R Final Project"
author: "Hannah Shapiro"
output:
  html_document:
    df_print: paged
---




```{r}
library(knitr)
library(tidyverse)
library(broom)
library(psych)
library(ggplot2)
library(ggpubr)
```

# Part 1

```{r}
# Importing the data sets without changing them from base yet

hcp.cognitive <- read.csv("~/Downloads/hcp-cognitive.csv")
hcp.demos <- read.csv("~/Downloads/hcp-demos.csv")
hcp.personality <- read.csv("~/Downloads/hcp-personality.csv")
```

```{r}
# Transforming the demographic data set

finalData<- hcp.demos
finalData$Age<- NULL
finalData$Gender<- NULL

for(i in 1:100) {
  if(!(is.na(hcp.demos$M_22.25[i])) | !(is.na(hcp.demos$M_31.35[i])) | !(is.na(hcp.demos$M_26.30[i])) | !(is.na(hcp.demos$M_36.[i]))) {
    finalData$Gender[i] <- "M"
  }
  else {
    finalData$Gender[i] <- "F"
  }
}
for(i in 1:100) {
if(!is.na(finalData$M_22.25[i]) | !is.na(finalData$F_22.25[i])) {
  finalData$Age[i] <- "22-25"
}
else if(!is.na(finalData$M_26.30[i]) | !is.na(finalData$F_26.30[i])) {
  finalData$Age[i] <- "26-30"
}
else if(!is.na(finalData$M_31.35[i]) | !is.na(finalData$F_31.35[i])) {
  finalData$Age[i] <- "31-35"
} 
else if(!is.na(finalData$M_36.[i]) | !is.na(finalData$F_36.[i])) {
  finalData$Age[i] <- "36+"
}   
else {
  finalData$Age[i] <- "NA"
}
}

demoTransform <- finalData[,c(1,10,11)]
demoTransform






```

```{r}

# Cognitive dataset transformation

finalData$CardSort_AgeAdj <- NULL
finalData$PMAT24_A_CR <- NULL
finalData$ProcSpeed_AgeAdj <- NULL
finalData$ListSort_AgeAdj <- NULL

for(i in 1:100) {
  if(is.na(hcp.cognitive$groupA.CardSort_AgeAdj[i])) {
    finalData$CardSort_AgeAdj[i] <- hcp.cognitive$groupB.CardSort_AgeAdj[i]
    finalData$PMAT24_A_CR[i] <- hcp.cognitive$groupB.PMAT24_A_CR[i]
    finalData$ProcSpeed_AgeAdj[i] <- hcp.cognitive$groupB.ProcSpeed_AgeAdj[i]
    finalData$ListSort_AgeAdj[i] <- hcp.cognitive$groupB.ListSort_AgeAdj[i]
}
else {
   finalData$CardSort_AgeAdj[i] <- hcp.cognitive$groupA.CardSort_AgeAdj[i]
   finalData$PMAT24_A_CR[i] <- hcp.cognitive$groupA.PMAT24_A_CR[i]
   finalData$ProcSpeed_AgeAdj[i] < -hcp.cognitive$groupA.ProcSpeed_AgeAdj[i]
   finalData$ListSort_AgeAdj[i] <- hcp.cognitive$groupA.ListSort_AgeAdj[i]
}
}
cogTransform <- finalData[,c(12,13,14,15)]

combineDemoCog <- cbind(demoTransform, cogTransform)

wrangled <- combineDemoCog[order(combineDemoCog$Subject, decreasing=T),]


orderedPersonality<-hcp.personality %>% 
  pivot_wider(names_from = Measure, values_from = Score)



transformedPersonality <- orderedPersonality[,-1]

hcp <- cbind(wrangled, transformedPersonality)
hcp

```

# Part 2

```{r, warning=FALSE}

agree <- ggplot(hcp, aes(NEOFAC_A)) + geom_density() + geom_label(aes(33,0.08, label = paste("Mean:", round(mean(NEOFAC_A, na.rm = T,), digits = 2)))) + geom_label(aes(33,0.01, label = paste("SD:", round(sd(NEOFAC_A, na.rm = T), digits = 2)))) + labs(x = "Agreeableness", y = "Density")

conscientious <- ggplot(hcp, aes(NEOFAC_C)) + geom_density() + geom_label(aes(35,0.08, label = paste("Mean:", round(mean(NEOFAC_C, na.rm = T,), digits = 2)))) + geom_label(aes(35,0.01, label = paste("SD:", round(sd(NEOFAC_C, na.rm = T), digits = 2)))) + labs(x = "Conscientiousness", y = "Density")

extraversion <- ggplot(hcp, aes(NEOFAC_E)) + geom_density() + geom_label(aes(33,0.08, label = paste("Mean:", round(mean(NEOFAC_E, na.rm = T,), digits = 2)))) + geom_label(aes(33,0.01, label = paste("SD:", round(sd(NEOFAC_E, na.rm = T), digits = 2)))) + labs(x = "Extraversion", y = "Density")

neuroticism <- ggplot(hcp, aes(NEOFAC_N)) + geom_density() + geom_label(aes(20,0.08, label = paste("Mean:", round(mean(NEOFAC_N, na.rm = T,), digits = 2)))) + geom_label(aes(20,0.01, label = paste("SD:", round(sd(NEOFAC_N, na.rm = T), digits = 2)))) + labs(x = "Neuroticism", y = "Density")

openness <- ggplot(hcp, aes(NEOFAC_O)) + geom_density() + geom_label(aes(33,0.08, label = paste("Mean:",round(mean(NEOFAC_O, na.rm = T,), digits = 2)))) + geom_label(aes(33,0.01, label = paste("SD:",round(sd(NEOFAC_O, na.rm = T), digits = 2)))) + labs(x = "Openness", y = "Density")

figure<-ggarrange(agree, conscientious, extraversion, neuroticism, openness)
annotate_figure(figure, bottom = text_grob("Figure 1: Distributions of the Big 5"))

```

While not perfectly normal, each aspect of the Big 5 does see a highest density in the middle, which then tapers off into the tails, similar to the normal distribution.

# Part 3

For my multivariate regression, I want to see whether or not life satisfaction is affected by 3 traits from the Big 5: extraversion, openness to experience, and neuroticism. I predict that life satisfaction will increase with extraversion and openness to experience, but will decrease with neuroticism.

```{r}
fit <- lm(LifeSatisf_Unadj ~ NEOFAC_E + NEOFAC_O + NEOFAC_N, data = hcp)
summary(fit)
```

Looking at these findings we can see that, holding all other predictors constant, extraversion, openness to experience, and neuroticism are significant predictors of overall life satisfaction. Namely life satisfaction has a positive relationship with extraversion, while it has a negative relationship with openness to experience and neuroticism.

The test yielded a F-statistic of 16.48, which also has a significant p-value at a testing level when alpha equals 0.05. Because of this signficance we know that a model with these 3 personality trait predictors is better than a model fitted without these traits.

In terms of fit, the model has a R-squared value of 0.3423 and a R-squared adjusted value of 0.3215. This means that there is a moderate correlation between the three predictors and life satisfaction in terms of how Cohen's d assesses effect size.

```{r}
fitResid <- augment(fit)
ggplot(fitResid, aes(x = .fitted, y = .resid)) + geom_point() + labs(x = "Fitted", y = "Residuals")

```

Overall, the residuals seem to be more skewed as the fitted values increase. This is not what we want to see in a residuals plot, as it means that there is something else influencing the regression that is making the residuals not be constant.


# Part 4

```{r}
summary(hcp$LifeSatisf_Unadj)
sd(hcp$LifeSatisf_Unadj)
life <- rnorm(99, 54.92, 10.11)

summary(hcp$NEOFAC_E)
sd(hcp$NEOFAC_E, na.rm = T)
extraversion <- rnorm(99, 30.98, 5.74)


summary(hcp$NEOFAC_O)
sd(hcp$NEOFAC_O, na.rm = T)
openness <- rnorm(99, 27.26, 6.61)


summary(hcp$NEOFAC_N)
sd(hcp$NEOFAC_N, na.rm = T)
neuroticism <- rnorm(99, 16.04, 7.67)


fitVec <- NULL
regression <- NULL
p <- NULL
sig <- NULL
for(i in 1:5000) {
  life <- rnorm(99, 54.92, 10.11)
  extraversion <- rnorm(99, 30.98, 5.74)
  openness <- rnorm(99, 27.26, 6.61)
  neuroticism <- rnorm(99, 16.04, 7.67)
  fitVec <- lm(life~extraversion+openness+neuroticism)
  regression <- anova(fitVec)
  p <- regression$`Pr(>F)`[1]
  if(p < 0.05) {
   sig[i] <- 1
 }
 else {
   sig[i] <- 0
 }
}

mean(sig)

```

Power is the probability that a hypothesis test finds an effect, given that there is a significant effect to be found. Because the power level when the simulated data is normal is `r mean(sig)`, we can tell that given a significant effect, we have a probability of `r mean(sig)` of finding it. This is extremely low, which leads me to believe that the original data is not normal enough to simulate it with normal data. 
